Analyzing Performance Bottleneck in Graph Neural Network Training: An Experimental View

### 2020-07-20 Manuscript/sec2-review-of-graph-neural-network.tex
修正gaan公式

Different GNNs adopt different kinds of graph neurons and have different definitions of the three functions.
$\phi$ and $\Sigma$ are the \emph{edge calculation} functions.
They are conducted over every edge in $\mathcal{G}$.
$\gamma$ is the \emph{vertex calculation} function.
It is conducted over every vertex in $\mathcal{G}$.
\tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} list the edge functions and the vertex functions of typical GNNs, respectively.
For ChebNet, we report its GNN sub-layer in the tables. 
A ChebNet layer consists of $K$ GNN sub-layers and a summation layer:
$\boldsymbol{H}^{l+1} = \sum_{k=1}^K{\boldsymbol{Z}^{(k)} \textcolor{blue}{\boldsymbol{W}^{(k)}}}$ with the GNN sub-layers $\boldsymbol{Z}^{(1)}=\boldsymbol{H}^l$, $\boldsymbol{Z}^{(2)}=\hat{\boldsymbol{L}}\boldsymbol{H}^l$, and $\boldsymbol{Z}^{(k)}=2\hat{\boldsymbol{L}}\boldsymbol{Z}^{(k-1)} - \boldsymbol{Z}^{(k-2)}$, where $\boldsymbol{H}^l$ is the matrix of output hidden feature vectors of the layer $l-1$.
For GAT, a GAT layer consists of two sub-layers and it conducts part of the vertex calculation before the two sub-layers.
For GaAN, a GaAN layer consists of four sub-layers: the first sub-layer calculates the summation $\sum_{k \in \mathcal{N}(j)}\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}}  \boldsymbol{h}_k^l] )$, and the other three layers calculate $\boldsymbol{m}^l_{j,i,1}$/$\boldsymbol{m}^l_{j,i,2}$/$\boldsymbol{m}^l_{j,i,3}$.

@@

						 & \alpha_{j, i, (k)} = \frac {\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l}_{i,(k)} \parallel \hat{\boldsymbol{h}}^{l}_{j,(k)}] ))} {\hat{\boldsymbol{a}}^{l,0}_{i,(k)}} \\
						 & \text{Multi-head concatenation}: \boldsymbol{m}_{j, i}^{l,1} = \parallel_{k=1}^K \delta(\alpha_{j, i, (k)} \hat{\boldsymbol{h}}^{l}_{j,(k)})                                                                                                                                             \\
						 & \text{Multi-head average}: \boldsymbol{m}_{j, i}^{l,1} = \frac{1}{K} \sum_{k=1}^K \delta(\alpha_{j, i, (k)} \hat{\boldsymbol{h}}^{l}_{j,(k)})

@@

						 & \alpha_{j, i}^k = \frac {\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l, k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}} \boldsymbol{h}_i^l] )} {\sum_{k \in \mathcal{N}(j)}\exp(\textcolor{blue}{\boldsymbol{a}^T} [ \textcolor{blue}{\boldsymbol{W}^{l,k}_{xa}} \boldsymbol{h}_j^l \parallel \textcolor{blue}{\boldsymbol{W}^{l,k}_{ya}}  \boldsymbol{h}_k^l] )} \\
						 & \boldsymbol{m}_{j, i, 1}^l = \parallel_{k=1}^K \delta(\alpha_{j, i}^k \textcolor{blue}{\boldsymbol{W}^{l,k}_v} \boldsymbol{h}_j^{l})                                                                                                                                                                                                                                                                                               \\
						 & \boldsymbol{m}_{j, i, 2}^l = \textcolor{blue}{\boldsymbol{W}^l_m} \boldsymbol{h}_j^{l}                                                                                                                                                                                                                                                                                                                                             \\
						 & \boldsymbol{m}_{j, i, 3}^l = \boldsymbol{h}_j^l

@@

@@

						 & \boldsymbol{g}_i = \textcolor{blue}{\boldsymbol{W}}^l_g  [\boldsymbol{h}_i^{l} \parallel \boldsymbol{s}_{i, 2}^l \parallel \boldsymbol{s}_{i, 3}^l] \\
						 & \boldsymbol{h}_i^{l+1} = \textcolor{blue}{\boldsymbol{W}}^l_o [\boldsymbol{h}_i^l \parallel (\boldsymbol{g}_{i} \odot \boldsymbol{s}_{i, 1}^l) ]

@@

		For Neural FPs, $\textcolor{blue}{\boldsymbol{W}}^{l, |\mathcal{N}(i)|}$ is the weight matrix for nodes with degree $|\mathcal{N}(i)|$ at layer $l$.}

@@

\textbf{GaAN} (high vertex \& edge calculation complexity):
Based on the multi-head machanism, GaAN introduces a convolutional subnetwork to control the weight of each head.
Thus, it introduces more edge-related parameters to train.
Moreover, GaAN aggregates three messages simutaneously, further increasing the edge calculational complexity.

>>>>

Different GNNs adopt different kinds of graph neurons and have different definitions of the three functions.
$\phi$ and $\Sigma$ are the \emph{edge calculation} functions.
They are conducted over every edge in $\mathcal{G}$.
$\gamma$ is the \emph{vertex calculation} function.
It is conducted over every vertex in $\mathcal{G}$.
\tablename~\ref{tab:gnn_overview_edge} and \tablename~\ref{tab:gnn_overview_vertex} list the edge functions and the vertex functions of typical GNNs, respectively.
For ChebNet, we report its GNN sub-layer in the tables. 
A ChebNet layer consists of $K$ GNN sub-layers and a summation layer:
$\boldsymbol{H}^{l+1} = \sum_{k=1}^K{\boldsymbol{Z}^{(k)} \textcolor{blue}{\boldsymbol{W}^{(k)}}}$ with the GNN sub-layers $\boldsymbol{Z}^{(1)}=\boldsymbol{H}^l$, $\boldsymbol{Z}^{(2)}=\hat{\boldsymbol{L}}\boldsymbol{H}^l$, and $\boldsymbol{Z}^{(k)}=2\hat{\boldsymbol{L}}\boldsymbol{Z}^{(k-1)} - \boldsymbol{Z}^{(k-2)}$, where $\boldsymbol{H}^l$ is the matrix of output hidden feature vectors of the layer $l-1$.
For GAT, a GAT layer consists of two sub-layers and it conducts part of the vertex calculation before the two sub-layers.
For GaAN, a GaAN layer consists of four sub-layers: the first sub-layer calculates the summation $\boldsymbol{m}_{j, i, (k)}^{l,0}  = \exp(\textcolor{blue}{FC_{\theta^{l, k}_{xa}}} (\boldsymbol{h}_j^l) \odot \textcolor{blue}{\boldsymbol{FC}_{\theta^{l,k}_{ya}}} (\boldsymbol{h}_i^l))$, and the other three layers calculate $\boldsymbol{m}^l_{j,i,1}$/$\boldsymbol{m}^l_{j,i,2}$/$\boldsymbol{m}^l_{j,i,3}$.

@@

						 & \alpha^l_{j, i, (k)} = \frac {\exp(LeakyReLU(\textcolor{blue}{\boldsymbol{a}^T} [\hat{\boldsymbol{h}}^{l}_{i,(k)} \parallel \hat{\boldsymbol{h}}^{l}_{j,(k)}] ))} {\hat{\boldsymbol{a}}^{l,0}_{i,(k)}} \\
						 & \text{Multi-head concatenation}: \boldsymbol{m}_{j, i}^{l,1} = \parallel_{k=1}^K \delta(\alpha^l_{j, i, (k)} \hat{\boldsymbol{h}}^{l}_{j,(k)})                                                                                                                                             \\
						 & \text{Multi-head average}: \boldsymbol{m}_{j, i}^{l,1} = \frac{1}{K} \sum_{k=1}^K \delta(\alpha^l_{j, i, (k)} \hat{\boldsymbol{h}}^{l}_{j,(k)})

@@

						 & \text{Sub-layer 0:} \\
						 & \boldsymbol{m}_{j, i, (k)}^{l,0}  = \exp(\textcolor{blue}{FC_{\theta^{l, k}_{xa}}} (\boldsymbol{h}_j^l) \odot \textcolor{blue}{\boldsymbol{FC}_{\theta^{l,k}_{ya}}} (\boldsymbol{h}_i^l)) \\
						 & \text{Sub-layer 1:} \\
						 & \alpha^l_{j, i, (k)} = \frac {\exp(\textcolor{blue}{FC_{\theta^{l, k}_{xa}}} (\boldsymbol{h}_j^l) \odot \textcolor{blue}{\boldsymbol{FC}_{\theta^{l,k}_{ya}}} (\boldsymbol{h}_i^l))} {\hat{\boldsymbol{a}}^{l, 0}_{i, (k)}}\\
						 & \boldsymbol{m}_{j, i}^{l, 1} = \parallel_{k=1}^K \alpha_{j, i, (k)} \textcolor{blue}{\boldsymbol{FC}^h_{\theta^l_v}} \hat{\boldsymbol{h}}_{j, (k)}^l \\                                                                                                                                                                                                                                                                                           
						 & \text{Sub-layer 2:} \\
						 & \boldsymbol{m}_{j, i}^{l, 2} = \textcolor{blue}{\boldsymbol{FC}_{\theta^l_m}} \boldsymbol{h}_j^{l}  \\
						 & \text{Sub-layer 3:} \\
						 & \boldsymbol{m}_{j, i}^{l, 3} = \boldsymbol{h}_j^l

@@

		Blue variables are model parameters to learn. In GaAN, $FC^{\alpha}_{\theta}(x) = \alpha(\boldsymbol{W}\boldsymbol{x} + \boldsymbol{b})$, 
		where $\theta=\{\boldsymbol{W}, \boldsymbol{b}\}$, $\theta$ with different subscripts mean different transformation parameters, we denote $h(\cdot)$
		to be the $LeakyReLU$ for activation functions and $FC_{\theta}(x)$ means applying no activation function after the linear transform.

@@

						& \text{Sub-layer 1,2,3:}\\
						& \boldsymbol{g}_i = \textcolor{blue}{\boldsymbol{FC}_{\theta^l_g}}(\boldsymbol{h}_i^{l} \parallel \boldsymbol{s}_{i}^{l, 2} \parallel \boldsymbol{s}_{i}^{l, 3}) \\
						& \boldsymbol{h}_i^{l+1} = \textcolor{blue}{\boldsymbol{FC}_{\theta^l_o}} (\boldsymbol{h}_i^l \parallel (\boldsymbol{g}_{i} \odot \boldsymbol{s}_{i}^{l, 1}))

@@

		For Neural FPs, $\textcolor{blue}{\boldsymbol{W}}^{l, |\mathcal{N}(i)|}$ is the weight matrix for vertices with degree $|\mathcal{N}(i)|$ at layer $l$. 
		In GaAN, $FC^{\alpha}_{\theta}(x) = \alpha(\boldsymbol{W}\boldsymbol{x} + \boldsymbol{b})$, 
		where $\theta=\{\boldsymbol{W}, \boldsymbol{b}\}$, $\theta$ with different subscripts mean different transformation parameters, we denote $h(\cdot)$
		to be the $LeakyReLU$ for activation functions and $FC_{\theta}(x)$ means applying no activation function after the linear transform.}

@@

\textbf{GaAN} (high vertex \& edge calculation complexity):
Based on the multi-head machanism, GaAN introduces a convolutional subnetwork to control the weight of each head. Different to multi-head machanism in GAT, GaAN project the center vertex feature to the query vector (dimension is $d_v$) and
the neighboring vertex features are projected to get key and value vectors (dimensions are $d_{xa}$ and $d_{ya}$, because they are set to the same value in the original paper experiment, we use $d_a$ to represent $d_{xa}$ and $d_{ya}$ in the description that follows).
$K$ is the number of attention heads. The weight of each head is calculated by a convolutional network, which takes the feature of each vertex and its neighbors as input to get the gated values with number of $K$. 
The convolutional network is built by combining average pooling and max pooling. Moreover, the max pooling needs to project the vertex feature to a transformed vector with dimension of $d_m$, costing an additional computational overhead.
Each GaAN layer consists of four sub-layers. The first sub-layer aims to get the summation of attention weights of every vertex in each head.
The second sub-layer uses to aggreate neighboring vertices's query vector with the normalized attention weights $\alpha_{j, i, (k)}$ for every head. 
And the third sub-layer is used in max-pooling and it gets the element-wise max vector of neighboring vertices' transformed vectors by message-passing.
The last sub-layer gets element-wise mean vector of neighboring vertices' vectors by message-passing. The gate values $\boldsymbol{g}^l_i$ is obtained by using a fully-connected layer with vector concatenating $\boldsymbol{h}^{l}_i$, $\boldsymbol{s}^{l, 2}_i$ and $\boldsymbol{s}^{l, 3}_i$.
At last, GaAN uses a fully-connected layer with concatenating $\boldsymbol{h}^{l}_i$ and the result of element-wise multiplying $\boldsymbol{g}^{l}_i$ and $\boldsymbol{s}^{l, 0}_i$ to get the output hidden vector.
